\ttitle

### 2013-07-03 Appendices/AppendixA-summaries.tex
wip

\subsection{Machine learning: the art and science of algorithms that make sense of data}
Book by Peter Flach: \cite{flach2012machine}.
Mainly about chapter 7, ``Linear Models''.
Most important: section 7.3 - 7.5, about support vector machines and non-linearity.
\textbf{Some parts are direct text; do not use this text directly!}

@@

\subsection{Change Point Detection In Time Series Data Using Support Vectors}
Paper by Fatih Camci \cite{camci2010change}
About segmentation with SVMs.
Will be main material for section \ref{svm} about SVMs.

@@

\subsection{Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance}
Carla Inclan and George C. Tiao \cite{inclan1994use}, 1944, 162 refs.

@@

\subsection{Change-Point Detection in Time-Series Data by Direct Density-Ratio Estimation}
Kawahara and Sugiyama \cite{kawahara2009change}, 2009, 45 refs.

@@

%--------------------------------------------
\subsubsection{Introduction}
The problem of change-point detection is well studied over the last decades in the field of statistics.
A common statistical formulation of change-point detection is to consider the probability distributions over past and present data intervals, and regard the target time as a change-point if the two distributions are significantly different.
Some approaches (such as \gls{cusum} and GLR) make use of the \emph{log-likelihood ratio}, and are extensively explored in the data mining community.
Many approaches (novelty detection, maximum-likelihood ratio, learning of autoregressive models, subspace identification) rely on pre-specified parametric models (probability density models, autoregressive models, state-space models).
That makes it less applicable to real-life problems.
There have been some non-parametric density estimation approaches proposed, but that is known to be a hard problem.
The key idea of this paper to directly estimate the \emph{ratio} of the probability densities (also known as \emph{importance}).
The Kullback-Leibler Importance Estimation Procedure (KLIEP) is an example, but it is a batch algorithm.
This paper introduces an online version of the KLIEP algorithm and develops a flexible and computationally efficient change-point detection method.
This method is equipped with a natural cross validation procedure and thus the value of tuning parameters can be objectively determined.

@@

>>>>

\subsection{Machine learning: the art and science of algorithms that make sense of data}
Book by Peter Flach:~\cite{flach2012machine}.
Mainly about chapter 7, ``Linear Models''.
Most important: section 7.3 - 7.5, about support vector machines and non-linearity.
\textbf{Some parts are direct text; do not use this text directly!}

@@

\subsection{Change Point Detection In Time Series Data Using Support Vectors}
Paper by Fatih Camci~\cite{camci2010change}
About segmentation with SVMs.
Will be main material for section \ref{svm} about SVMs.

@@

\subsection{Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance}
Carla Inclan and George C. Tiao~\cite{inclan1994use}, 1944, 162 refs.

@@

\subsection{Change-Point Detection in Time-Series Data by Direct Density-Ratio Estimation}
Kawahara and Sugiyama~\cite{kawahara2009change}, 2009, 45 refs.

@@

%--------------------------------------------
\subsubsection{Introduction}
The problem of change-point detection is well studied over the last decades in the field of statistics.
A common statistical formulation of change-point detection is to consider the probability distributions over past and present data intervals, and regard the target time as a change-point if the two distributions are significantly different.
Some approaches (such as \gls{cusum} and GLR) make use of the \emph{log-likelihood ratio}, and are extensively explored in the data mining community.
Many approaches (novelty detection, maximum-likelihood ratio, learning of autoregressive models, subspace identification) rely on pre-specified parametric models (probability density models, autoregressive models, state-space models).
That makes it less applicable to real-life problems.
There have been some non-parametric density estimation approaches proposed, but that is known to be a hard problem.
The key idea of this paper to directly estimate the \emph{ratio} of the probability densities (also known as \emph{importance}).
The \gls{kliep} is an example, but it is a batch algorithm.
This paper introduces an online version of the \gls{kliep} algorithm and develops a flexible and computationally efficient change-point detection method.
This method is equipped with a natural cross validation procedure and thus the value of tuning parameters can be objectively determined.

@@

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{Outlier Detection methods}
\subsection{A Survey of Outlier Detection Methodologies}
Hodge and Austin,~\cite{hodge2004survey}, 2004, 734 refs.

\subsection{Summary}
This survey tooks a look at different methodologies that perform outlier detection.
It gives two definitions of outliers, one of which relates to the problem statement in this thesis (taken from~\cite{barnett1994outliers}):
\begin{quote}An observation (or subset of observations) which appears to be inconsistent with the remainder of that set of data.\end{quote}

The survey introduces three fundamental approaches to the problem of outlier detection:
\begin{description}
  \item[Type 1] - no prior knowledge about the data; analogous to \emph{unsupervised clustering}. Flags the remote points as outliers; mainly batch-processing systems.
  \item[Type 2] - model (and requires) both normal and abnormal data; analogous to \emph{supervised classification}. Is able to process new data online.
  \item[Type 3] - Model only normal data (and in very few cases abnormal). Generally refered to as \emph{novelty detection}, analogous to \emph{semi-supervised recognition or detection}. It only requires pre-classified normal data. It aims to define a boundary of normality.
\end{description}

The type of approaches of interest in this research is of \textbf{type 3}.

### 2013-07-03 Appendices/AppendixB-camci.tex
wip

In \cite{liu2013change} it is stated that the \emph{Support Vector Machine} by Cortes and Vapnik \cite{cortes1995support} is representative example of this principle.
Instead of estimating the more general data generating probability distributions, it only learns a decision boundary to differentiate between the two distributions.

The proposed method \gls{svdd} of Tax and Duin \cite{tax1999support, tax2004support} models the boundary of data under consideration.
Thereby it characterizes a data set and can be used to detect novel data or outliers.
The performance is compared to methods which model the distribution's density instead, using Receiver-Operating Characteristic (ROC) curves and false negative rates.
The compared methods are:
\begin{inparaenum}[\itshape (1)]
  \item normal density which estimates the mean and covariance matrix;
  \item the Parzen density where the width of the Parzen kernel is estimated;
  \item a Gaussian Mixture Model optimized using EM; and
  \item the Nearest-Neighbor Method which compares the local density of an object with the density of the nearest neighbor in the target set (and is thus, just as SVDD, a boundary-based method).
\end{inparaenum}
The results show that when the problem formulation is to characterize an area in a feature space (and not the complete density distribution) SVDD gives a good data description.

The study of Tax on one-class classifiers \cite{tax2001one} further compares density methods, boundary methods (amongst which SVDD) and reconstruction methods.
One promesing result for SVDD is that it performs well on read-world data, for which generalization is needed.

@@

The paper of Yin \etal \cite{yin2008sensor} makes a distinction between similarity based (using a defined distance measure) and model based (which characterize the data using predictive models) approach.
The one-class SVM is used in the model-approach to filter out normal activities in order to detect abnormality behavior.

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{Change point definition}
The method of Camci \cite{camci2010change} regards a change point as the moment in time that the underlying stochastic process has changed, say from $p^1$ to $p^2$.
It assumes that each of these stochastic processes is modeled following a Gaussian distribution, such that a change can occur in the value of the mean and/or the variance; $p^1 \sim N(\mu_1, \sigma_1^2)$.
The CUSUM-based method of \cite{inclan1994use} also regards each semgent as a Gaussian distribution.

The method if Kawahara \etal \cite{kawahara2009change} is based on the log likelihood ratio of test samples, and the method by Liu \etal \cite{liu2013change} uses a comparable dissimilarity measure using the KLIEP algorithm.

The method of Chamroukhi \etal \cite{chamroukhi2013joint} is based on a Hidden Markov Model and logistic regression.
It assumes a $K$-state hidden process with a (hidden) state sequence, each state providing the parameters (amongst which the order) for a polynomial.
The order of the model segment is determined by model selecting, often using the Bayesian Information Criterion (BIC) or the similair Akaike Information Criterion (AIC) \cite{akaike1974new}, as in \cite{he2008activity}.

@@

Why use RBF/Gaussian kernel? Is it because of one-class SVM of because of form the data? Why not polynomial/linear?
\begin{itemize}
  \item ``Use RBF when relation between class and data is non-linear''
  \item ``RBF uses less paramters (C for penalty/soft margin and gamme for kernel width) than non-linear polynomial kernels''
  \item (arguments from \cite{zhang2011optimal} ``Optimal model selection for posture recognition in home-based healthcare'')
  \item Survey \cite{hodge2004survey} states RBF is similair to Gaussian Mixture model (page 25).
\end{itemize}

>>>>

In \cite{liu2013change} it is stated that the \gls{svm} by Cortes and Vapnik \cite{cortes1995support} is a representative example of this principle.
Instead of estimating the more general data generating probability distributions, it only learns a decision boundary to differentiate between the two distributions.

The proposed method \gls{svdd} of Tax and Duin \cite{tax1999support, tax2004support} models the boundary of data under consideration.
Thereby it characterizes a data set and can be used to detect novel data or outliers.
The performance is compared to methods which model the distribution's density instead, using \gls{roc} curves and false negative rates.
The compared methods are:
\begin{inparaenum}[\itshape (1)]
  \item normal density which estimates the mean and covariance matrix;
  \item the Parzen density where the width of the Parzen kernel is estimated;
  \item a Gaussian Mixture Model optimized using EM; and
  \item the Nearest-Neighbor Method which compares the local density of an object with the density of the nearest neighbor in the target set (and is thus, just as \gls{svdd}, a boundary-based method).
\end{inparaenum}
The results show that when the problem formulation is to characterize an area in a feature space (and not the complete density distribution) \gls{svdd} gives a good data description.

The study of Tax on \glspl{occ} \cite{tax2001one} further compares density methods, boundary methods (amongst which \gls{svdd}) and reconstruction methods.
One promesing result for \gls{svdd} is that it performs well on read-world data, for which generalization is needed.

@@

The paper of Yin \etal \cite{yin2008sensor} makes a distinction between similarity based (using a defined distance measure) and model based (which characterize the data using predictive models) approach.
The \gls{oc-svm} is used in the model-approach to filter out normal activities in order to detect abnormality behavior.

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{Change point definition}
The method of Camci \cite{camci2010change} regards a change point as the moment in time that the underlying stochastic process has changed, say from $p^1$ to $p^2$.
It assumes that each of these stochastic processes is modeled following a Gaussian distribution, such that a change can occur in the value of the mean and/or the variance; $p^1 \sim N(\mu_1, \sigma_1^2)$.
The \gls{cusum}-based method of \cite{inclan1994use} also regards each semgent as a Gaussian distribution.

The method if Kawahara \etal \cite{kawahara2009change} is based on the log likelihood ratio of test samples, and the method by Liu \etal \cite{liu2013change} uses a comparable dissimilarity measure using the \gls{kliep} algorithm.

The method of Chamroukhi \etal \cite{chamroukhi2013joint} is based on a Hidden Markov Model and logistic regression.
It assumes a $K$-state hidden process with a (hidden) state sequence, each state providing the parameters (amongst which the order) for a polynomial.
The order of the model segment is determined by model selecting, often using the \gls{bic} or the similair \gls{aic} \cite{akaike1974new}, as in \cite{he2008activity}.

@@

Why use RBF/Gaussian kernel? Is it because of \gls{oc-svm} of because of form the data? Why not polynomial/linear?
\begin{itemize}
  \item ``Use RBF when relation between class and data is non-linear''
  \item ``RBF uses less paramters (C for penalty/soft margin and gamme for kernel width) than non-linear polynomial kernels''
  \item (arguments from \cite{zhang2011optimal} ``Optimal model selection for posture recognition in home-based healthcare'')
  \item Survey \cite{hodge2004survey} states RBF is similair to Gaussian Mixture model (page 25).
\end{itemize}

### 2013-07-03 Chapters/Chapter2/density_ratio.tex
wip

Many approaches to detect change points, regarded as a change in the underlying probabilistic generation, monitor the logarithm of the likelihood ratio between two consecutive intervals.
Some methods which rely on this are novelty detection, maximum-likelihood estimation and online learning of autoregressive models \cite{kawahara2009change}.
A limitation of these methods is that they rely on pre-specified parametric models.
Non-parametric models for density estimation have been proposed, but it is said to be a hard problem \cite{hardle2004nonparametric, sugiyama2012density}.
A solutions to this is to estimate the \emph{ratio} of probabilities instead of the probabilities themselves.
One of the recent methods to achieve this is the Kullback-Leibler Importance Estimation Procedure (KLIEP) by Sugiyama \etal \cite{sugiyama2008direct}.

The method proposed by Kawahara and Sugiyama \cite{kawahara2009change} is composed of an online version of the KLIEP algorithm.
The method also considers \emph{sequences} of samples (rather than samples directly) because the time series samples are generally not independent over time.
An advantage over other non-parametric approaches, such as sequential one-class support vector machines, is that the model has an natural cross-validation procedure.
This makes that the value of tuning parameters, such as the kernel bandwidth, can be objectively obtained.

@@

The procedure of the method proposed by Kawahara and Sugiyama \cite{kawahara2009change} is to first apply the batch KLIEP algorithm with model selection for initial parameter $\vectorsym{\alpha}$ and kernel width calculation.
Then for every new sample the reference and test intervals are shifted and the calculated parameters $\vectorsym{\alpha}$ are updated.
Finally the logarithm of the likelihood ratio is evaluated.
If it is beyond the threshold $\mu$ the current time is reported as a change point.

>>>>

Many approaches to detect change points, regarded as a change in the underlying probabilistic generation, monitor the logarithm of the likelihood ratio between two consecutive intervals.
Some methods which rely on this are novelty detection, maximum-likelihood estimation and online learning of autoregressive models \cite{kawahara2009change}.
A limitation of these methods is that they rely on pre-specified parametric models.
Non-parametric models for density estimation have been proposed, but it is said to be a hard problem \cite{hardle2004nonparametric, sugiyama2012density}.
A solutions to this is to estimate the \emph{ratio} of probabilities instead of the probabilities themselves.
One of the recent methods to achieve this is the \gls{kliep} by Sugiyama \etal \cite{sugiyama2008direct}.

The method proposed by Kawahara and Sugiyama \cite{kawahara2009change} is composed of an online version of the \gls{kliep} algorithm.
The method also considers \emph{sequences} of samples (rather than samples directly) because the time series samples are generally not independent over time.
An advantage over other non-parametric approaches, such as sequential one-class support vector machines, is that the model has an natural cross-validation procedure.
This makes that the value of tuning parameters, such as the kernel bandwidth, can be objectively obtained.

@@

The procedure of the method proposed by Kawahara and Sugiyama \cite{kawahara2009change} is to first apply the batch \gls{kliep} algorithm with model selection for initial parameter $\vectorsym{\alpha}$ and kernel width calculation.
Then for every new sample the reference and test intervals are shifted and the calculated parameters $\vectorsym{\alpha}$ are updated.
Finally the logarithm of the likelihood ratio is evaluated.
If it is beyond the threshold $\mu$ the current time is reported as a change point.
